---
title: "Fisheriescape scores"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fisheriescape scores}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```
# Overview
<p style="font-size: 16px;">
This vignette generates weekly fishery-specific and cumulative fisheriescape scores from the site scores and CEU scores generated in the previous vignettes.
</p>
```{r echo=FALSE, warning = FALSE, message = FALSE, error=FALSE}
library(knitr)
library(kableExtra)
```


# R packages
## CRAN R packages
```{r libraries, warning = FALSE, message = FALSE, error=FALSE}
library(dplyr)
library(terra)
library(ggplot2)
library(tidyterra)
```

## GitHub R packages
The following packages are all available from https://jolenesutton.github.io.
```{r gitlibraries, warning = FALSE, message = FALSE, error=FALSE}
library(fisheriescape) 
```



# Get site score and CEU scores
```{r results='hold'}
df.ss<-fs.data.site.score
df.ceu<-fs.data.ceu

dim(df.ss)
dim(df.ceu)

head(df.ss)
head(df.ceu)
```


# Calculate fisheriescape scores
Fisheriescape is calculated as the product of the site score and CEU.
```{r echo=TRUE, message=FALSE}

df.fs<-left_join(df.ss,df.ceu)


df.fs$fs<-df.fs$ss*df.fs$ceu

# Because spatial grid cells can be associated with multiple fishing areas we need to sum fs for each grid cell for each fishery and sw
# We no longer need to group by NAFO division
df.fs<-df.fs|>
  group_by(fishery,gear.name,sw,GRID_ID)|>
  summarize(fs=sum(fs))


summary(df.fs)

```

```{r results='hold',echo=FALSE}
kable(head(df.fs), 
      caption = "head(df.fs)",format = "html")|>
  kable_styling(full_width = FALSE) |>
  column_spec(1:ncol(df.fs), width = "300px", extra_css = "white-space: nowrap;")|>
  scroll_box(width = "100%")
```
