---
title: "Fisheriescape vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fisheriescape vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup,  warning = FALSE, message = FALSE, error=FALSE}
library(fisheriescape)
library(gslSpatial)
library(terra)
library(ggplot2)
library(tidyterra)
library(dplyr)
```


# Load example files and a 10 x 10 km2 hexagonal grid
```{r, warning = FALSE, message = FALSE}
df<-crab
fa.poly<-terra::vect(system.file("extdata/sample.fa.poly.shp", package="fisheriescape"))

grid<-gslSpatial::get_shapefile('hex')
grid<-terra::crop(grid,ext(fa.poly)*1.05)
```


# Site score
```{r, warning = FALSE, message = FALSE}
df.ss<-distinct_at(df,vars('event.id'),.keep_all = T)
df.ss<-df.ss[-which(is.na(df.ss$x)|is.na(df.ss$y)),]

ggplot()+
  geom_spatvector(data=fa.poly,fill='white')+
  geom_spatvector(data=grid,fill=NA)+
  geom_point(data=df.ss,aes(x,y))
```


```{r, warning = FALSE, message = FALSE}

ss<-suppressMessages(fs_site_score(df.ss,fa.poly,grid=grid))# be patient
head(ss)
summary(ss)

index<-which(grid$GRID_ID%in%unlist(ss[which(is.na(ss$sum.av.count.fleet)),'GRID_ID']))

ggplot()+
  geom_spatvector(data=fa.poly,fill='white')+
  geom_spatvector(data=grid,fill=NA)+
  geom_spatvector(data=grid[index,],fill='red')#+geom_point(data=df.ss,aes(x,y))
# red hexes had data points, but they fall outside of the fa.poly, and so are not included in the site score calculations

ss<-ss[-which(is.na(ss$prop.count.fleet)),]
ss$site.score<-ss$prop.count.fleet
```


# Pivot tables

## Select distinct rows
```{r, warning = FALSE, message = FALSE}
df2<-distinct_at(df,vars('event.id','fleet','cfv','dateland','ctchdate',
                           'gear.amount',
                           'hours.fished'),.keep_all = T) 

df2<-droplevels(df2)
names(df2)
nrow(df2)
```

## Pivot 1
Function `fs_pivot1` summarizes data by variables trip.id and fleet. The resulting table contains the summed gear, maximum hours (if applicable), and mean days (if applicable).
```{r, warning = FALSE, message = FALSE}
pivot1<-fs_pivot1(df2,
      keep.cols = c('cfv','year','sw'),
       gear.col='gear.amount',
       hour.col='hours.fished')
pivot1
summary(pivot1) 
```

Now fill in the missing gear.amount with group averages.The user needs to choose the order of variables to group by that is most appropriate for the fishery of interest. By the end of this step, there should be no missing gear.amount. 
```{r, warning = FALSE, message = FALSE}
if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('cfv','year','sw'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('cfv','year'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('cfv'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('year','sw'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('year'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  print('There are still NAs after filling gear according to current steps.')}
```

There are no longer NAs in gear.amount. If object pivot1 contained NAs in hours.fished, we would repeat the above step for hours.fished. However, we can see that there are no NAs in hours.fished.
```{r, warning = FALSE, message = FALSE}
summary(pivot1) 
```


## Pivot 2
Function `fs_pivot2` summarizes data by user-specified groups, which must include year, standard week (sw), and either cfv or licence, and which should typically also include fleet. 

When trap.fishery=='yes', the resulting table contains the maximum gear, maximum hours (if applicable), and mean days (if applicable). 

When trap.fishery!='yes', the resulting table contains the mean gear, mean hours (if applicable), and mean days (if applicable).
```{r, warning = FALSE, message = FALSE}
pivot2<-fs_pivot2(pivot1,
                  trap.fishery='yes',
                  group.cols=c('cfv','fleet','year','sw'))
head(as.data.frame(pivot2))
summary(pivot2) 
```


## Pivot 3
Function `fs_pivot3` summarizes data by user-specified groups, which must include year and standard week (sw), and which should typically also include fleet. **Grouping should not include either cfv or licence.** A parameter, "id.col", needs to be specified as either cfv or licence.

When trap.fishery=='yes', the resulting table contains the number of distinct cfvs or licences ("n.vessels"), summed gear ("total.gear"), maximum hours ("soak.time", if applicable), and mean days ("days", if applicable).

When trap.fishery!='yes', the resulting table contains the number of distinct cfvs or licences ("n.vessels"), summed gear ("total.gear"), mean hours ("soak.time", if applicable), and mean days ("days", if applicable).

```{r, warning = FALSE, message = FALSE}
pivot3<-fs_pivot3(pivot2,
                  group.cols=c('fleet','year','sw'),
                  id.col='cfv',
                  trap.fishery='yes')
head(as.data.frame(pivot3))
summary(pivot3) 

```

## Seasons and Pivot 4
The function `fs_pivot4` is a little different from the previous pivot functions, as it performs a number of steps. It does not yet have a lot of flexibility, and needs to be used with caution. Don't use this function for fisheries that need information to be summarized by province in addition to by fleet.

**need to describe what function does, and resulting output**
```{r, warning = FALSE, message = FALSE}
pivot4<-fs_pivot4(df=df2,
                  trap.fishery='yes',
          pivot3, #output from fs_pivot3
          fleet.col='fleet', #should match between df and pivot3
          prov.col=NULL,
          area.note.col=NULL)
head(as.data.frame(pivot4))
summary(pivot4) 
```


# Common effort unit (CEU)
For most fisheries, use function `fs_calc_ceu` and follow the prompts. This function cannot yet be applied to American lobster, NAFO 4T.

**need to describe what function does for each fishery, and resulting output**
```{r, eval=FALSE}
ceu<-fs_calc_ceu(pivot4)
```

To simplify things for this vignette, we will use a pre-made CEU file.
```{r, eval=TRUE, warning = FALSE, message = FALSE}
ceu<-ceu
head(ceu)
```


# Fisheriescape

## Join site score with CEU and calculate fisheriescape
```{r, warning = FALSE, message = FALSE}
df<-left_join(ss,ceu)

df$fs<-df$site.score*df$av.ceu 

head(df)
summary(df) # a small number of missing fs or av.ceu are expected


#Checks
tapply(df$fs,df$fleet,summary)

```

## Plot
```{r, warning = FALSE, message = FALSE}
hex<-grid
naf<-get_shapefile('nafo.clipped')
naf<-project(naf,fa.poly)
names(naf)[1]<-'nafo'

#fa.poly<-gslSpatial::assign_polygons(fa.poly[,'fleet'],naf[,'nafo'])
#fa.poly<-fa.poly[which(fa.poly$nafo=='4T'),]

coast<-gslSpatial::get_shapefile('coastline')
coast<-project(coast,fa.poly)

WEEKS<-sort(unique(df$sw))
fig.save='no'  # option to save file to specified path. If not 'yes', the figures will not be save and will only appear in the graphics window. If 'yes', the figures will be saved to the specified path, but will not be printed in the graphics window.
path.results<-'....' #path to store figures if fig.save=='yes'

for(i in 1:length(WEEKS)){
  dat<-df[which(df$sw==WEEKS[i]),]
  index<-which(is.na(dat$fs))
  if(length(index)>0){dat<-dat[-index,]}
  
  if(nrow(dat)>0){
    hex.out<-hex[which(hex$GRID_ID%in%unique(dat$GRID_ID)),]
    hex.out<-merge(hex.out,dat)
    
    pl<-ggplot()+
      geom_spatvector(data=fa.poly,fill='white')+
          geom_spatvector(data=coast,fill='grey')+
      geom_spatvector(data=hex.out,aes(fill=fs),col=NA)+
      scale_fill_viridis_c(direction=1,
                           option='turbo',
                           name="Fisheriescape",
                           limits = c(min(df$fs,na.rm=TRUE), max(df$fs,na.rm=TRUE)))+
      coord_sf(crs='epsg:4269',datum='epsg:4269',xlim=c(-69,-60),ylim=c(45.5,49.75))+
      ggtitle(paste0('SW_',WEEKS[i]))+
      theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank())
    
   if(fig.save=='no'){print(pl)}else{ggsave(pl, 
                          file = paste0(path.results,'FS_SW', WEEKS[i], '.jpg', sep = ''), 
                          units = 'in', dpi = 600,height=4,width=6)}
  }
}
```

