---
title: "Fisheriescape vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fisheriescape vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup,  warning = FALSE, message = FALSE, error=FALSE}
library(fisheriescape)
library(gslSpatial)
library(terra)
library(ggplot2)
library(tidyterra)
library(dplyr)
```


# Load example files and a 10 x 10 km2 hexagonal grid
```{r, warning = FALSE, message = FALSE}
df<-crab
fa.poly<-terra::vect(system.file("extdata/sample.fa.poly.shp", package="fisheriescape"))

grid<-gslSpatial::get_shapefile('hex')
grid<-terra::crop(grid,ext(fa.poly)*1.05)
```


# Site score
First, select distinct rows by variable 'event.id'.
```{r, warning = FALSE, message = FALSE}
df.ss<-distinct_at(df,vars('event.id'),.keep_all = T)
df.ss<-df.ss[-which(is.na(df.ss$x)|is.na(df.ss$y)),]

ggplot()+
  geom_spatvector(data=fa.poly,fill='white')+
  geom_spatvector(data=grid,fill=NA)+
  geom_point(data=df.ss,aes(x,y))
```


Next, use function `fs_site_score`. This function first counts the number of fishing events per grid cell (hexagon) by standard week and year. Then it calculates the mean number of fishing events by standard week for each grid cell. Finally, it calculates the **site score** of each grid cell by dividing the mean number of fishing events in each grid cell by the mean number of fishing events in each grid cell summed by fleet and standard week.
```{r, warning = FALSE, message = FALSE,eval=FALSE}
ss<-suppressMessages(fs_site_score(df.ss,fa.poly,grid=grid))# be patient
head(ss)
summary(ss)
```

```{r, warning = FALSE, message = FALSE,echo=FALSE}
ss<-suppressMessages(fs_site_score(df.ss,fa.poly,grid=grid))# be patient
names(ss)[which(names(ss)=='prop.count.fleet')]<-'site.score'
head(ss)
summary(ss)
```


In the figure below, red grid cells had data points, but these points fall outside of the fishing area polygons, and so are not included in the site score calculations.
```{r, warning = FALSE, message = FALSE, fig.show="hold"}
index<-which(grid$GRID_ID%in%unlist(ss[which(is.na(ss$sum.av.count.fleet)),'GRID_ID']))

ggplot()+
  geom_spatvector(data=fa.poly,fill='white')+
  geom_spatvector(data=grid,fill=NA)+
  geom_spatvector(data=grid[index,],fill='red')

ss<-ss[-which(is.na(ss$site.score)),]
```


# Pivot tables (pre-CEU steps)

## Select distinct rows
```{r, warning = FALSE, message = FALSE}
df2<-distinct_at(df,vars('event.id','fleet','cfv','dateland','ctchdate',
                           'gear.amount',
                           'hours.fished'),.keep_all = T) 

df2<-droplevels(df2)
names(df2)
nrow(df2)
```

## Pivot 1
Function `fs_pivot1` summarizes data by variables trip.id and fleet. The resulting table contains the summed gear, maximum hours (if applicable), and mean days (if applicable).
```{r, warning = FALSE, message = FALSE}
pivot1<-fs_pivot1(df2,
      keep.cols = c('cfv','year','sw'),
       gear.col='gear.amount',
       hour.col='hours.fished')
pivot1
summary(pivot1) 
```

Now fill in the missing gear.amount with group averages.The user needs to choose the order of variables to group by that will be most appropriate for the fishery of interest. By the end of this step, there should be no missing gear.amount. 
```{r, warning = FALSE, message = FALSE}
if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('cfv','year','sw'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('cfv','year'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('cfv'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('year','sw'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  pivot1<-fisheriescape::fs_fill_col(pivot1,group.cols=c('year'),update.col = 'sum.gear',fun=mean)}

if(length(which(is.na(pivot1$sum.gear)))>0){
  print('There are still NAs after filling gear according to current steps.')}
```

There are no longer NAs in gear.amount. If object pivot1 contained NAs in hours.fished, we would repeat the above step for hours.fished. However, we can see that there are no NAs in hours.fished.
```{r, warning = FALSE, message = FALSE}
summary(pivot1) 
```


## Pivot 2
Function `fs_pivot2` summarizes data by user-specified groups, which must include year, standard week (sw), and either cfv or licence, and which should typically also include fleet. 

When trap.fishery=='yes', the resulting table contains the maximum gear, maximum hours (if applicable), and mean days (if applicable). 

When trap.fishery!='yes', the resulting table contains the mean gear, mean hours (if applicable), and mean days (if applicable).
```{r, warning = FALSE, message = FALSE}
pivot2<-fs_pivot2(pivot1,
                  trap.fishery='yes',
                  group.cols=c('cfv','fleet','year','sw'))
head(as.data.frame(pivot2))
summary(pivot2) 
```


## Pivot 3
Function `fs_pivot3` summarizes data by user-specified groups, which must include year and standard week (sw), and which should typically also include fleet. **Grouping should not include either cfv or licence.** A parameter, "id.col", needs to be specified as either cfv or licence.

When trap.fishery=='yes', the resulting table contains the number of distinct cfvs or licences ("n.vessels"), summed gear ("total.gear"), maximum hours ("soak.time", if applicable), and mean days ("days", if applicable).

When trap.fishery!='yes', the resulting table contains the number of distinct cfvs or licences ("n.vessels"), summed gear ("total.gear"), mean hours ("soak.time", if applicable), and mean days ("days", if applicable).

```{r, warning = FALSE, message = FALSE}
pivot3<-fs_pivot3(pivot2,
                  group.cols=c('fleet','year','sw'),
                  id.col='cfv',
                  trap.fishery='yes')
head(as.data.frame(pivot3))
summary(pivot3) 

```

## Seasons and Pivot 4
The function `fs_pivot4` is a little different from the previous "pivot" functions, as it performs a number of steps. It does not yet have a lot of flexibility, and needs to be used with caution. Don't use this function for fisheries that need information to be summarized by province in addition to by fleet.

Function `fs_pivot4` first identifies the start and end of the fishing season for each year and fleet, and calculates the proportion of each week that fishing occurs. It then joins the proportion of week fished to the output from function `fs_pivot3`. 
```{r, warning = FALSE, message = FALSE}
pivot4<-fs_pivot4(df=df2,
                  trap.fishery='yes',
          pivot3, #output from fs_pivot3
          fleet.col='fleet', #should match between df and pivot3
          prov.col=NULL,
          area.note.col=NULL)
head(as.data.frame(pivot4))
summary(pivot4) 
```


# Common effort unit (CEU)
For most fisheries, use function `fs_calc_ceu` with the output from `fs_pivot4` and follow the prompts. This function cannot yet be applied to American lobster, NAFO 4T.

For snow crab, `fs_calc_ceu` uses the following steps to calculate CEU:

  1. rope.soak.time = soak.time/24
  1. magnitude = total.gear
  1. intensity = magnitude X rope.soak.time
  1. pre.ceu = intensity X proportion.week.fished
  1. ceu = sum pre.ceu by fleet and standard week, then divide by the number of years examined
```{r, eval=FALSE}
df.ceu<-fs_calc_ceu(pivot4)

head(df.ceu)
```

```{r, echo=FALSE, eval=TRUE, warning = FALSE, message = FALSE}
df.ceu<-ceu
names(df.ceu)[3]<-'ceu'
head(df.ceu)
```


# Fisheriescape
The **fisheriescape** is the product of the **site score** multiplied by the **CEU**.

## Join site score with CEU and calculate fisheriescape
```{r, warning = FALSE, message = FALSE}
df<-left_join(ss,df.ceu)

df$fs<-df$site.score*df$ceu 

head(df)
summary(df) # a small number of missing fs or av.ceu are expected


#Checks
tapply(df$fs,df$fleet,summary)

```

## Plot
```{r, warning = FALSE, message = FALSE,eval=FALSE}
hex<-grid
naf<-get_shapefile('nafo.clipped')
naf<-project(naf,fa.poly)
names(naf)[1]<-'nafo'

coast<-gslSpatial::get_shapefile('coastline')
coast<-project(coast,fa.poly)

WEEKS<-sort(unique(df$sw))
fig.save='no'  # option to save file to specified path. If not 'yes', the figures will not be save and will only appear in the graphics window. If 'yes', the figures will be saved to the specified path, but will not be printed in the graphics window.
path.results<-'....' #path to store figures if fig.save=='yes'

for(i in 1:length(WEEKS)){
  dat<-df[which(df$sw==WEEKS[i]),]
  index<-which(is.na(dat$fs))
  if(length(index)>0){dat<-dat[-index,]}
  
  if(nrow(dat)>0){
    hex.out<-hex[which(hex$GRID_ID%in%unique(dat$GRID_ID)),]
    hex.out<-merge(hex.out,dat)
    
    pl<-ggplot()+
      geom_spatvector(data=fa.poly,fill='white')+
          geom_spatvector(data=coast,fill='grey')+
      geom_spatvector(data=hex.out,aes(fill=fs),col=NA)+
      scale_fill_viridis_c(direction=1,
                           option='turbo',
                           name="Fisheriescape",
                           limits = c(min(df$fs,na.rm=TRUE), max(df$fs,na.rm=TRUE)))+
      coord_sf(crs='epsg:4269',datum='epsg:4269',xlim=c(-69,-60),ylim=c(45.5,49.75))+
      ggtitle(paste0('SW_',WEEKS[i]))+
      theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank())
    
   if(fig.save=='no'){print(pl)}else{ggsave(pl, 
                          file = paste0(path.results,'FS_SW', WEEKS[i], '.jpg', sep = ''), 
                          units = 'in', dpi = 600,height=4,width=6)}
  }
}
```

